\documentclass[12pt]{article}
\title{CASS Agenda}
\author{Patryk Kozlowski}
\date{\today}
\begin{document}
\maketitle
\section{medical}
\subsection{my situation}
How much did the outgoing accessibility specialist tell you about my medical situation? If it wasn't much, then I am happy to fill you in with some details. I am going to the hospital to get an infusion once a month, and am just taking oral chemotherapy at home, which I will be continuing until May of 2024.
\subsection{student wellness services}
I will need monthly lab draws for oncology purposes and then also a weekly subcutaneous shot. The Caltech medical director, back in February told me that you guys would have such abilities. How should I go about arranging these things?
\section{non-medical}
\subsection{room}
I will be living in Bechtel 164a. Would there be an opportunity for my mother and me to come visit sometime during this summer before I move in? Or is somebody living there for summer research? We live in Irvine, which is about 1 hour away and we would be willing to drive up. I am able to use a standing desk, since I am dictating, and I also have a treadmill at home that I use every day, so we are just interested in what can fit in my dorm room.
\subsection{board}
I use an assistive device in order to walk, so I would be interested in walking to the cafeteria for some meals, but somebody would have to help me with serving the food. Previously, when I spoke with Mark over a year ago, he said that you guys could do things like bring food to a certain place where I would eat. What could my cafeteria accommodations look like? And who should I contact to work them out?
\subsection{talon}
I am voice coding and doing any kind of text generation, like here, using dictation. I am just curious what the use of Talon looks like in a rigorous academic setting, like Caltech. I have asked around on the Talon Slack page, and they advised me that it's best to do everything remotely, which is fairly obvious. Are you working with any students who are using Talon? If so, would you be able to connect me with them, so I could ask them a few questions about their experience with in-person learning at Caltech?
\end{document}
