\documentclass[12pt]{article}
\title{Dining Accommodations}
\author{Patryk Kozlowski}
\date{\today}
\begin{document}
\maketitle
\section{Overnight Oats for Breakfast}
For over the past year, I have made this a routine that I would like to continue at Caltech. Because of the steroid treatment that I get as part of my chemotherapy, I have a larger appetite than I did before the illness, particularly in the morning. It is nice to have something to eat when I get up in the morning, and I think this would be fairly easy to continue at Caltech. My parents will help me organize a mini fridge at my dorm at Caltech, so it would just be a matter if someone from the dining services could prepare a few servings for me in advance. I have been preparing them 5 days in advance at home where I store them in the fridge, and this doesn't have to exactly continue at Caltech, but I would like to continue the basic idea of having my breakfast ready in the morning when I wake up. Perhaps, it would be also possible for someone from the dining services to bring me some sort of fruit, like an apple, with my overnight oats, so that I can take it in my backpack and eat it the next day before lunch?

\section{Cafeteria Accommodations for Lunch and Dinner}
For lunch and dinner meals, I would like to come to the cafeteria to have some social interaction. However, I will need some accommodations relating to my physical mobility. I use an assistive device, like a walker, and so I need to have both hands on my assistive device at all times, meaning that I cannot carry a tray to a table. If I recall correctly, the Caltech dining services prepare a menu for the week in advance. I was thinking that I would be able to pick what I want to eat based on this pre-prepared menu, and then when I come to the cafeteria, someone would just help me bring the food to a table. I am able to use utensils just fine, so dining accommodations will primarily be a mobility accommodation for me.

\section{Previous Experience?}
This is my first time dealing with such dining accommodations. Do you guys have any pertinent experience in the past for students with similar needs to mine? I am just not sure if I am thinking of everything that I should be thinking about.
\end{document}
