\documentclass[12pt]{article}
\usepackage{amsmath}
\usepackage{physics}
\usepackage{graphicx}
\usepackage{hyperref}
\usepackage{listings} % Required for insertion of code
\usepackage{xcolor} % Required for custom colors

% Define custom colors
\definecolor{codegreen}{rgb}{0,0.6,0}
\definecolor{codegray}{rgb}{0.5,0.5,0.5}
\definecolor{codepurple}{rgb}{0.58,0,0.82}
\definecolor{backcolour}{rgb}{0.95,0.95,0.92}

% Setup the style for code listings
\lstdefinestyle{mystyle}{
    backgroundcolor=\color{backcolour},   
    commentstyle=\color{codegreen},
    keywordstyle=\color{magenta},
    numberstyle=\tiny\color{codegray},
    stringstyle=\color{codepurple},
    basicstyle=\ttfamily\footnotesize,
    breakatwhitespace=false,         
    breaklines=true,                 
    captionpos=b,                    
    keepspaces=true,                 
    numbers=left,                    
    numbersep=5pt,                  
    showspaces=false,                
    showstringspaces=false,
    showtabs=false,                  
    tabsize=2
}

% Activate the style
\lstset{style=mystyle}

\author{Patryk Kozlowski}
\title{ssa}
\date{\today}
\begin{document}
\maketitle
Form SSA-3373 (10-2020)
Page 4 of 10
SECTION C - INFORMATION ABOUT DAILY ACTIVITIES
6. Describe what you do from the time you wake up until going to bed.
7. Do you take care of anyone else such as a wire/husband, children, grandchildren,
Yes
No
parents, friend, other?
If "YES," for whom do you care, and what do you do for them?
8. Do you take care of pets or other animals?
Yes
No
If "YES," what do you do for them?
9. Does anyone help you care for other people or animals?
Yes
No
If "YES," who helps, and what do they do to help?
10. What were you able to do before your illnesses, injuries, or conditions that you can't do now?
11. Do the illnesses, injuries, or conditions affect your sleep?
Yes\\
No\\
If "YES," how?\\
12. PERSONAL CARE (Check here if NO PROBLEM with personal care.)\\
a. Explain how your illnesses, injuries, or conditions affect your ability to:\\
\\
Dress\\
\\
Bathe\\
\\
Care for hair\\\\
Not applicable\\\\
Shave\\
\\
Feed self\\
\\
Use the toilet\\\\
I need flushable wipes or a bidet\\\\
Form SSA-3373 (10-2020)\\
Page 5 of 10\\
b. Do you need any special reminders to take care of personal needs and groomi\\ng?\\
\\
No\\\\
If "YES," what type of help or reminders are needed?\\\\
c. Do you need help or reminders taking medicine?\\\\
No\\
If "YES," what kind of help do you need?\\
13. MEALS\\
a. Do you prepare your own meals?\\
No\\
If "Yes," what kind of food do you prepare? (For example, sandwiches, frozen d\\inners, or complete meals with several courses.)
\\
How often do you prepare food or meals? (For example, daily, weekly, monthly.)\\
\\
How long does it take you?\\
Any changes in cooking habits since the illness, injuries, or conditions began\\?
I transitioned to a plant based diet
b. If "No," explain why you cannot or do not prepare meals.\\
My balance is impaired for standing and my fine motor function is impaired for working with my hands
14. HOUSE AND YARD WORK\\
a. List household chores, both indoors and outdoors, that you are able to do. \\(For example, cleaning, laundry, household repairs, ironing, mowing, etc.)
Other people help me for most of the chores, but I am able to do laundry myself by kneeling down
b. How much time does it take you, and how often do you do each of these thing\\s?
I only do my laundry around once every two weeks, but fording it takes quite a bit of time because of my fine motor impairment
c. Do you need help or encouragement doing these things?\\
No\\
If "YES," what help is needed?\\
d. If you don't do house or yard work, explain why not.\\
Because of the impairments to my balance and fine motor function
10024082\\
Form SSA-3373 (10-2020)\\
Page 6 of 10\\
15. GETTING AROUND\\
a. How often do you go outside?\\
Every day to ago to class
If you don't go out at all, explain why not.\\
b. When going out, how do you travel? (Check all that apply.)\\
Walk\\
Ride in a car\\
c. When going out, can you go out alone?\\
Yes\\
If "NO," explain why you can't go out alone.\\
d. Do you drive?\\
No\\
\\
If you don't drive, explain why not.\\
My fine mortar function and reflexes are impaired.
16. SHOPPING\\
a. If you do any shopping, do you shop: (Check all that apply.)\\
By computer\\
b. Describe what you shop for.\\
Just personal items
c. How often do you shop and how long does it take?\\
I mostly just do it of of an application on my smart phone, so it doesn't take much time
17. MONEY\\
a. Are you able to:\\
Pay bills\\
Yes\\
Count change\\
Yes\\
Handle a savings account\\
Yes\\
Use a checkbook/money orders\\
Yes\\
Explain all "NO" answers.\\
b. Has your ability to handle money changed since the illnesses, injuries, or \\conditions began?
No\\
If "YES," explain how the ability to handle money has changed.\\
10024082\\
Form SSA-3373 (10-2020)\\
Page 7 of 10\\
18. HOBBIES AND INTERESTS\\
a. What are your hobbies and interests? (For example, reading, watching TV, se\\wing, playing sports, etc.)\\
I like watching political comedy on my phone and going for long walks with my family and friends
b. How often and how well do you do these things?\\
I go for wax every day, but I need an assistive device, like a walker
c. Describe any changes in these activities since the illnesses, injuries, or \\conditions began.\\
Since my immune system is compromise from chimo, I am not able to connect with peers like I used to
19. SOCIAL ACTIVITIES\\
a. Do you spend time with others? (Check all that apply.)\\
In person On the phone Email Texting Mail\\
Video Chat (for example Skype or Facetime) \\
b. Describe the kinds of things you do with others.\\
Going for walks or working on homeworks with other students at school
\\
How often do you do these things?\\
Every day
c. List the places you go on a regular basis. (For example, church, community \\center, sports events, social groups, etc.)
\\
Classroom for college lectures
Do you need to be reminded to go places?\\
no
How often do you go and how much do you take part?\\
During the weekdays; I will ask the professor questions occasionally
Do you need someone to accompany you?\\
If "YES", explain.\\
No\\
d. Do you have any problems getting along with family, friends, neighbors, or \\others?
No\\
If "YES," explain.\\
e. Describe any changes in social activities since the illnesses, injuries, or\\ conditions began.
10024082\\
| Lifting | Walking | Stair Climbing | |\\
| :---: | :---: | :---: | :---: |\\
| Squatting | | | |\\
| Bending | | | |\\
| Standing | Talking | Completing Tasks | Getting Along With Others |\\
| Reaching | | |  |\\
For this one, it might be helpful to call and discuss
Form SSA-3373 (10-2020)
Page 8 of 10
SECTION D - INFORMATION ABOUT ABILITIES
20. a. Check any of the following items that your illnesses, injuries, or conditions affect:
\begin{tabular}{|c|c|c|c|}
\hline Lifting & Walking & Stair Climbing & Understanding \\
\hline Squatting & Sitting & Seeing & Following Instructions \\
\hline Bending & Kneeling & Memory & Using Hands \\
\hline Standing & Talking & Completing Tasks & Getting Along With Others \\
\hline Reaching & Hearing & Concentration & \\
\hline
\end{tabular}

Please explain how your illnesses, injuries, or conditions affect each of the items you checked. (For example, you can only lift [how many pounds], or you can only walk [how far].)
b. Are you: Right Handed?\\
c. How far can you walk before needing to stop and rest?
About an hour\\
If you have to rest, how long before you can resume walking?
5 min\\
d. For how long can you pay attention?
This function was not impaired by my neurological event\\
e. Do you finish what you start? (For example, a conversation, chores, reading, watching a movie.)
Yes

f. How well do you follow written instructions? (For example, a recipe.)
Very well\\
g. How well do you follow spoken instructions?
Very well\\
h. How well do you get along with authority figures? (For example, police, bosses, landlords or teachers.)
Very well\\
i. Have you ever been fired or laid off from a job because of problems getting along with other people?
no\\
If "YES," please explain.

If "YES," please give name of employer.
10024082
Form SSA-3373 (10-2020)\\
Page 9 of 10\\
j. How well do you handle stress?\\
Very well\\
k. How well do you handle changes in routine?\\
Not very well I need to have the appropriate ADA physical accommodations most of rn
I. Have you noticed any unusual behavior or fears?\\
\\
No\\
If "YES," please explain.\\
21. Do you use any of the following? (Check all that apply.)\\
Walker\\
Wheelchair\\
Which of these were prescribed by a doctor?\\
The wheel chair
When was it prescribed?\\
11/21???
\\
When do you need to use these aids?\\
 Both inside the house and outdoors
10024082\\
\end{document}
