\documentclass[12pt,letterpaper]{article}
\usepackage[american]{babel}
\usepackage{csquotes}
\usepackage[style=mla,backend=biber]{biblatex}
\usepackage[letterpaper]{geometry}
\usepackage{times}
\geometry{top=1.0in, bottom=1.0in, left=1.0in, right=1.0in}
\usepackage{lipsum}  

%Doublespacing
\usepackage{setspace}
\doublespacing

%Rotating tables (e.g. sideways when too long)
\usepackage{rotating}

%Fancy-header package to modify header/page numbering (Put your surname in the surname field)
\usepackage{fancyhdr}
\pagestyle{fancy}
\lhead{} 
\chead{} 
\rhead{Kozlowski \thepage}
\lfoot{} 
\cfoot{} 
\rfoot{} 
\renewcommand{\headrulewidth}{0pt} 
\renewcommand{\footrulewidth}{0pt} 
%To make sure we actually have header 0.5in away from top edge
%12pt is one-sixth of an inch. Subtract this from 0.5in to get headsep value
\setlength\headsep{0.333in}

%You put your name title and professor name in here 
\renewcommand{\maketitle}{\makemlaheader}

%You put your name title and professor name in here 
\renewcommand{\maketitle}{\makemlaheader}
\newcommand{\C}{\autocite}

\newcommand{\makemlaheader}{
Patryk Kozlowski \\
Eve Darian-Smith\\
IS 17\\
\begin{center}\textnormal{Final Paper}\end{center}

% for some reason, this blank line is necessary
}

\addbibresource{citations.bib} % you would use your own bib file here
\begin{document}
\begin{flushleft}

%%%%First page name, class, etc
%%%%Changes paragraph indentation to 0.5in
\setlength{\parindent}{0.5in}

\maketitle
There is scientific consensus regarding the fact of anthropogenic climate change due to fossil fuel consumption.According to report by the united nations enter national candle on climate change, there is growing evidence of climate action, but it is not enough to reduce global heading to the target of1.5C. to try and reach this decarbonization gas, there has been a renewed attention on nuclear fission and energy sources, which have a long history, dating back to the disaster that have been incur noble in the 19 ads to the melt down of the Fukushima nuclear plant in the 2010. The particular issue attend, which regards the controlled removable of radioactive wastewater into the Japanese ocean, seems like fusional issue, but in fact it is very international. This paper will start by examining the Fukushima waste water cases and then will explore the status of nuclear fission and energy on the global stage, particularly in Germany, Which was undergoing a face out of nuclear fission and energy before the Russian invasion of Ukraine, but defect that natural has supplied by Russia, which was formerly seen as a transition fossil hull,. Is cleaner, but is no longer a viable source of energy, due to geopolitical ramifications. It will be argued thought if there is a desire to reach the united nations decarbonization goes, humanity must accept nuclear fission and energy as an essential part of the energy makes, as other renewable and energy sources become cheaper. The current issue in japan regarding the waste water is not just a regional one, but it is rather important globally to, as being extremely net picky about how this waste water is removed, merely reenforces an attitude which is set against nuclear vision and energy among the public, which is dangerous as humanity grapples with fossil hug given primate change.

First, I should address the criticism to japan's plan to release the Fukushima wastewater into the ocean. On the raged and of the spectrum, neighboring geopolitical rival china is accusing japan of using their common ocean as its private sower dump pivot other other, and more reasonable part of the spectrum, human rights and environmental activists argue that not enough research has been done to justify dumping the waste water into the ocean yet. It is true that water is treated like commodity that is taken for granted and when it is tainted, this has proven to have devastating consequences. A famous, domestic example of this was the recent water curses infant, Michigan. In what was largely a economically dominated move, the mostly white municipality leaders decided to switch the water source from the Detroit water system to the Flint river, which was known to be contaminated with that. However, the sum levities between sent and japan and their. The flat river was known to be contaminated with led, and white leaders overlooked this element at the expense of the largely black citizenry, in a common case of virtual injustice tied to environmental injustice. In the case of japan, there is a third party watchdog involved kama namely the international atomic anergy agency kama which has deemed the plan a safe. More generally, however, this negative attitude Of the global community towards nuclear fission anergy is dangerous and will be explored in the following Pecos.

 in particular, Germany, which had formerly used nuclear fission energy as a vital part of its energy mix. closed down most of its reactors. There was some debate about back sharking on this action after the natural gama supplied by Russia kama was no longer available after its invasion of Ukraine kama but nevertheless the german government has continued. The green party in Germany was a chief supporter of shutting down nuclear fission capacity since the twentieth century. Back then, the, crisis was not so amide, and nuclear fission generation seemed like more oven environmental concern. However, things have changed since then. In the wake of shedding down its nuclear fission capacity, the german government must instead rely on call for its electricity generation. however, it is erotic, because science has reached a consensus regarding the safety of nuclear fission generation versus cole. "furthermore, There is large public concern about nuclear fission generation being attributed to cancer, but studies have repeatedly shown that this is false, with fossil holds actually presenting a greater public health concern. Nuclear efficient generation is something that humanity has shown a capacity to safely control, so it presents a tactical source of energy generation, while other renewables well start the become more practical in the future. Additionally, there had been recent advances and actinide cemetery for the efficient processing of waste relating to nuclear fission generation. While the scientific community recognizes the suffering that the disasters in tenable and Fukushima have caused, there is a continued commandment two safe, civilian nuclear power. Instead, there has been significant interest in a technology called geo and gene ring, which involves purposely spring chemicals into the atmosphere in order to troy and call it. However, there is not much evidence supporting the viability of this strategy, boras there is a tactical solution right before our eyes in nuclear fission energy generation. Although, geo and gendering is different from nuclear fission generation in that it does not propose to actually generate energy, it represents a scheme to actually do nothing about the climate crisis right now and instead delay it, one we have do tools in nuclear fission and energy generation available to us write now.Humanity has shown a willingness to try and escape the climate crisis by researching de and renewing solutions and starting to accept equal socialism. However, these are long term solutions and they argument as practical as nuclear fission generation, which we know works, and should be instead focussed on.

There is overview of bill gates recent book on climate change, in which he talks a little bit about nuclear fission energy. The author takes issue with  bill gates oversimplifying the issue of nuclear fission and energy by reducing issues with dec analogy to demand of direct accidence it causes, and alsothe fact that gates comments on this issue, as an owner of accompany developing advanced nuclear fission reactors. while, it is true that nuclear fission and energy generation does have for reaching and volume mental impacts, the fact is that "The deaths from nuclear power (deaths per terawatt hour) are far eclipsed by its competitors with only 0.03 deaths per terawatt hour, compared to 32.72 for Brown Coal.  furthermore, it should be considered what nuclear fission generation has already done for us: "Over the past 50 years, nuclear energy reduced CO2 emissions by 60 gigatons- nearly two years’ worth of global energy emissions."The scientific community is working hard to make nuclear power a safe virality  and this technology must be leveraged in order to fervent ecological collapse.

\newpage

\printbibliography


\end{flushleft}
\end{document}
