\documentclass{article}
\usepackage[style=mla,backend=biber]{biblatex}
\addbibresource{fondle.bib}
\usepackage{hyperref}
\begin{document}
there is scientific consensus regarding the fact of anthropogenic climate change due to fossil fuel consumption. a transition towards decarbonization was occuring before Russia's invasion of Ukraine in 2022, but there has been a perverse of such policies since. the paper will examine the ramifications of the Ukraine consult from the .ens of neighboring Poland. the focus will be on the false fuel of hole which has along his survey Poland. first, the transition away from ko will be detailed prior to invasion of yukking, including the establishment of a silesian transformation fund by the european union. then it would be examined howthe combination of the version invasion with the election of an autocratic government Poland, has caused Poland to fall back on ko. the tradition of coining mining Poland has been used as a justification for why the polish autocratic government is unwilling to reach its climate golds, with the facing out of core mining. 
the salvation transformation, funded by the just transition program of the european union in 2022, added a social contract for ca miners transitioning away from work/ among other things, it pays for retirement pension benefits and the fit trending of these workers. former call miners find reemployment in a fa variety of industrious, ranging from heavy labor, like construction, to private property andsecurity affirms.
pivot following the invasion in Ukraine, Poland is signaling that it wants to use cote generate electricity past 2049, to bolster its energy security vivid this goes against the decarbonization Gauls of the european union, which have an ambitious plan of reaching a net zero in emissions by 2050. natural cast, which was typically supplied by Russia, is typically seen as a sort of transition away from the most polluting fossil fuel, which is coal, towards more renewable energy sources.
whole mining has a rich his survey in Poland. coal minders commend a level of respect, just below that of university professors medical professionals, and policemen. in addition, call mining is intergenerational. pervious can be seen here between polish call miners and yukking call miners, who are a source of national prided, as they are powering the country's resistance against the Russian invasion. furthermore, the call miners unions have history for participating actively in the solidarity movement of the 910 a dis, on the polish activist la valance was awarded the noble peace prize in 1983. as so, they have "ave enjoyed benefits that are the envy of their working class countrymen: An annual bonus of two months pay regardless of performance, company-sponsored holidays, retirement before 50, and no weekend shifts."
the police confirmed is using this tradition as a sort of justification for desiring with the european union, which it has publically conflicted with regarding its proposed judicial reforms of 2019. 
under the guise of preserving the tradition of core mining, the authoritarian regime in Poland that was put into power after the version invasion has hethere distinct features proposed by debian smith: a fraction of cooperative multiverses, anti environmentalism, and ultranationalism. the first point is evident, by the polish confirmed footing regularly with the governing european union, over its proposed jute shivery overhaul of 2019, among others. search a feuding with the Berean union is also a homework of the latter two features.
the second point is evidenced by the polish goverment continued financing of fosse fuse in coal. the first point here is more salta: the so called long tradition of cool mining is simply a keyword for ultranationalism. the polish nation is generally a unitary ethnic group, devoid of only minorities kama but the Salesian region where this coal mining takes place kama is like a Harland in Poland: kind of like how the state of west virginia, which is composed of the coal mining inthe story primarily, its the darling of american romantic ideology about what a white nationalism in America should look like. call mining also happens to be a labor intensive profession, and so it is typically an activity done by males, which further plays into the patriarchal portion of the white nationalist negative.
is just a test \C{darian-smith2019}
\end{document}