\documentclass[12pt]{article}
\title{Accommodations}
\author{Patryk Kozlowski}
\date{\today}
\begin{document}
\maketitle
\section{Lectures}
\subsection{Glean}
As you can see in my accommodations letter, there is mention of audio recordings of lectures. Glean essentially makes an audio recording of your lecture, so that I can go back to it later and review. Since I cannot handwrite, I want to try out this assistive tool in conjunction with handwritten notes that hopefully a classmate will share with me.
\subsection{Accessibility}
I use a walker, so stairs are a no-go for me, and doors that don't have an electric press are tough. I looked over the campus accessibility map and bridge, where your lectures are, seems fine. Could you just make sure that one of your recitation sections is in an accessible lecture hall too? I am thinking of using Glean in conjunction with volunteer handwritten notes for the recitation sections too, so I hope none of your teaching assistants have a problem with audio recordings.
\section{Problem Sets}
\subsection{Extensions}
There might be times where I am not feeling a hundred percent due to chemo, so as my accommodation letter says, would it be possible if I got 3 day extensions for 3 of my problem sets?
\subsection{Submission}
I will be completing all my sets in LaTeX anyway, so would it be possible to circumvent printing my problem set and submitting it manually? That is, would it be possible to just email the PDF to the teaching assistant responsible for grading that problem set?
\section{Exams}
\subsection{Extra Time}
I am using voice coding/GitHub Copilot to do LaTeX, but it would be my first time doing it in a timed setting. I understand that you don't want to give too much time for the exam, so that the students don't have way too much time to think about the problems. However, in my case, doing the derivations for the exam problems might take a lot of time as a motor skill. What do you propose here? I was thinking, prior to the midterm, I would have some experience using voice coding/GitHub Copilot to do the problem sets, so I could just give you an update about how much time I think I should need for the midterm.
\subsection{Submission}
Would it also be okay to do electronic submission, as with the problem sets?
\section{Questions}
Are there any other things that you think I should be thinking about that I haven't mentioned yet?
\end{document}