\documentclass[12pt]{article}
\usepackage{hyperref}
\title{Agenda}
\author{Patryk Kozlowski}
\date{\today}
\begin{document}
\maketitle
\section{Course Work}
What is your opinion regarding the necessity of me learning C at this stage of my education? A class which offers a light introduction to the language is being offered next term, so that's why I'm asking. I know Rui uses these concepts heavily for the Cuda implementation, but do your other students?
\section{What to do next summer?}
I likely will not be going on a vacation and will instead spend time at my home in Irvine where all of my doctors are. I want to discuss educational things I could do based in my home in Irvine.
\subsection{Site on starter project}
I recently went on the chanlab Github page, and I saw that there is something like a silicon starter project made by Matthew O'Rourke in 2019\\ \url{https://github.com/gkclab/tickle}\\
It seems like this might be kind of useful to me. Have any of your previous students done it? I know that this is a tricky question, but do you think it would be worth my time? It seems that no one has touched this in quite a while.
\subsection{Remote summer school}
Is there some kind of summer school that would be worth applying to, in electronic structure or CS?
\subsection{Striking up a connection at UCI}
This is the most exciting option since it would involve some in-person component. However, I will be graduating in the middle of June and then getting ready to move for whatever graduate program at the beginning of August; I am not sure whether I will be in a position to get seriously involved in a research project in this short time. I recall that you once said you know some people at UCI. Is there anyone you think it would be worth reaching out to?
\section{Recommendations}
The majority will be due in late November. Then, there will be the one for kings which is due in early march. I will remind you at the appropriate times.
\end{document}