\documentclass[12pt]{article}
\usepackage{physics}
\usepackage{amsmath}

\title{Agenda}
\author{Patryk Kozlowski}
\date{\today}

\begin{document}

\maketitle

\section{FCI Starter Project}
Quiz me!

\section{What to Do Next?} 

\subsection{Context}

\subsubsection{Senior Thesis}
Between now and the start of this school year, do you recommend that I do some short-term project or look into some technical code, or start on the reading that I imagine would be the starting point for my senior thesis? I have been in infrequent contact with the former undergraduate student of yours, Alessio, who started a senior thesis with you, and he said that he started with reading on things like embedding, quantum Monte Carlo, and FCI. These do seem like topics that I want to learn about if I want to pursue electronic structure in graduate school, so do you recommend similar readings for me to start the senior thesis or something different?

% \subsubsection{NSF Graduate Research Fellowship Program?}
% I don't know how much you know about the application, but there is a personal statement portion that I feel very comfortable with, and then a research plan section that I am kind of lost about. This is perfectly natural though because I have not been involved in current quantum chemistry research this past summer. In your opinion, is it worth applying now when I have a lot of time on my hands, or is it worth waiting until I matriculate into graduate school in order to apply and have a better idea about my research plan then?



\section{Recommendations}
I want to make writing recommendations for me as easy as possible for you. Would it be helpful to create a recommendation guide with anecdotes for you to look at or do you generally know what you want to write for me?

\section{Thank You for Suggesting Graduate Programs for Me}
Your recommendations gave me a lot of good options to think about, so I really appreciate it!

\end{document}
