\documentclass[margin,line]{resume}
\usepackage[latin1]{inputenc}
\usepackage[english,french]{babel}
\usepackage[T1]{fontenc}
\usepackage{fontawesome}
\usepackage{graphicx,wrapfig}
\usepackage{url}
\usepackage[colorlinks=true, pdfstartview=FitV, linkcolor=blue, citecolor=blue, urlcolor=blue]{hyperref}
\usepackage{verbatim}
\pdfcompresslevel=9


\begin{document}{\Large \textbf{\scshape Patryk Kozlowski}
\normalsize
\hspace{70mm}(949)-394-9220 $\bullet$ \url{pkozlows@caltech.edu}
}
\begin{resume}

% === EDUCATION ===
\section{\mysidestyle Education}
\textbf{California Institute of Technology}, Pasadena, CA\hfill June 2024 (expected)\\
Unweighted GPA: 3.91\\
BS; Chemistry

% === tag loco experience ===
\section{\mysidestyle Technical\\ Skills}
% show example of how to insert italicized text
\textbf{Programming:} Python with \emph{Numpy} (intermediate), PySCF (beginner), \LaTeX  (Advanced)

% === EXPERIENCE ===
\section{\mysidestyle Relevant Experience}
{\sl Research Fellow} \hfill November 2022 - Present\\
\textbf{California Institute of Technology}
\begin{itemize}
\item Work on quantum chemistry project advised by Prof. Garnet Chan
\item Implement and optimize Full CI for an $H_{6}$ chain (\url{https://github.com/pkozlows/fci})

\end{itemize}

{\sl Teaching Assistant} \hfill October 2020 - December 2020\\
\textbf{California Institute of Technology}
\begin{itemize}
\item Worked 9 hours/week as a TA for chemistry introductory QM course
\item Apart from grading responsibilities, held weekly recitations and office hours
\end{itemize}
{\sl John Stauffer Summer Undergraduate Research Fellow} \hfill June 2020 - September 2020 \\
\textbf{California Institute of Technology}
\begin{itemize}
\item Used quantum chemistry methods to run simulations with Prof. Garnet Chan
\item Computed surface energies of platinum using coupled-cluster theory
\item \emph{Publications and Presentations}
    \begin{itemize}
        \item Kozlowski, P. 2020. "Elucidating Catalysis with the "Gold Standard" of Quantum Chemistry". Oral session presented \href{https://youtu.be/pcNnGM0bYRw}{virtually} at Annual Caltech Fall SURF Seminar Day, October 17.
        \item Kozlowski, P. T. 2021. "\href{https://curj.caltech.edu/2021/06/29/elucidating-catalysis-with-the-gold-standard-of-quantum-chemistry/}{Elucidating Catalysis with the "Gold Standard" of Quantum Chemistry}". Caltech Undergraduate Research Journal, 21 (1).
    \end{itemize}
\end{itemize} 
{\sl John Stauffer Summer Undergraduate Research Fellow} \hfill June 2019 - September 2019 \\
\textbf{California Institute of Technology}
\begin{itemize}
\item Conducted physical inorganic chemistry research with Prof. Ryan Hadt
\item Developed a computational model for spin-phonon coupling in Co(III)
coordination complexes
\item \emph{Publications and Presentations}
    \begin{itemize}
            \item Kozlowski, P. 2019. "Spin-Phonon Coupling in Transition Metal Complexes." Oral session presented at Annual Caltech Fall SURF Seminar Day, October 19, Pasadena, CA.
            \item Higdon, N. J., A. T. Barth, P.T. Kozlowski, and R. G. Hadt. "\href{https://doi.org/10.1063/5.0006361}{Spin-Phonon Coupling and Dynamic Zero-Field Splitting Contributions to Spin Conversion Processes in Iron(II) Complexes}." Journal of Chemical Physics, 152 (20), 204306.
    \end{itemize}
\end{itemize}
\section{\mysidestyle Scholarships and Awards}
\begin{itemize}
    \item  \textnormal{Goldwater Scholar in Mathematics, Science, and Engineering}\hfill 2021
    \item \textnormal{Perpall Speaking Competition} Semifinalist, California Institute of Technology \hfill 2020
    \item \textnormal{Polish National Alliance Scholarship} \hfill 2020, 2021, 2023
    \item \textnormal{John Kopczynski Scholarship}, Polish University Club of Los Angeles \hfill 2019, 2020, 2023
    \item \textnormal{Richard Gorecki Scholarship}, Polish-American Congress \hfill 2020, 2023
    \item \textnormal{National Merit Scholar} \hfill 2018-2020, 2023
\end{itemize}

\section{\mysidestyle Other Experience}
{\sl Student-Faculty Conference Comittee Member}\hfill October 2020 - January 2021\\
\textbf{California Institute of Technology}
\begin{itemize}
\item Published an online survey to get feedback from students on the chemistry major
\item Discussed proposed curriculum changes with faculty
\end{itemize}
{\sl Social Director}\hfill October 2019 - December 2020\\
\textbf{Caltech Chemistry Club, Pasadena, CA}
\begin{itemize}
\item Organized monthly professional, community outreach, and social events
\item Recruited a distinguished chemical researcher for the club's annual speaker event
\end{itemize}
\newpage
{\sl Volunteer}\hfill June 2019 - July 2020\\
\textbf{California Institute of Technology}
\begin{itemize}
\item Held weekly tutoring sessions for the Caltech Y's RISE Program
\item Prepared underprivileged high school students for their STEM classes and the SAT
\end{itemize}
{\sl Athlete} \hfill January 2014 - December 2020\\
\textbf{California Institute of Technology}
\begin{itemize}
\item Started for Caltech's NCAA intercollegiate tennis team (ranked in the Top 25)
\end{itemize}

\section{\mysidestyle Languages}
\textbf{Polish: }\textnormal{Fluent}




\end{resume}   
\end{document}