\documentclass[12pt]{article}
\title{Agenda}
\author{Patryk Kozlowski}
\date{\today}
\begin{document}
\maketitle
\section{More about my medical situation}
\subsection{Chemo treatment}
My treatment runs until the beginning of May 2024. Up until that time, I am just going to the infusion center at the hospital once a month, and otherwise, I am just taking oral chemo at home every day. \emph{To summarize, once I graduate in June of 2024, I will no longer be under chemo treatment.}
\subsection{Stroke recovery prognosis}
Recovering from a stroke just depends on how many repetitions of what you choose to do every day. Since I am working hard on my physical and speech recovery, the suggested prognosis for a full recovery of these faculties is in five years. It has been over 2 years since my stroke.
\subsection{Questions}
I am pretty much an open book with regards to sharing details about my medical situation. Do you have any other questions that would be pertinent in deciding what fellowships and study abroad programs I can apply to?
\section{Graduate fellowship}
\subsection{Main criterion is academic excellence/achievement}
I imagine this includes the PD Soros and Hertz fellowships. I am pretty comfortable in applying to those as I am confident in my academic track record, and also I have no problem thinking of and writing essays over this summer.
\subsection{Research Proposal Is a Criterion}
I imagine that this includes NSF, DOE, and DOD. I am not as comfortable applying to this because I am not so confident about my ability to write a research proposal right now. However, in your previous email to me, you encouraged me to apply to this. Maybe part of the reason that I am hesitant to apply to these is because I took a quick glance at the NSF, saw that it required writing a research proposal, and then didn't look into it much more. Do you suggest that I give things a closer look? As you may have gathered from my Soros career essay, I only started doing the research that I want to do in graduate school a few months ago when I discovered dictation, so I am not at the stage to write a research proposal yet. But perhaps I will be come late summer or early fall.

\section{Study Abroad}
Being able to study abroad in England for a year or two before starting graduate school in the States would be an awesome experience, as you probably know. I am very much interested in applying to such a program.

\subsection{deferring Graduate Acceptances and Fellowships}
I probably need to do more research into how this works, but I imagine there is not a problem with deferring if I am doing a study abroad program for just a year or two?

\subsection{Study Abroad Programs that Look at Leadership as a Criterion}
I have not looked at this too closely yet, but for the leadership reason, I believe that the Marshall, gates-Cambridge, and Rhodes probably won't be a good fit for me. Do you agree, or do you suggest that I look into it more closely?

\subsection{Study Abroad Programs that Require Collaborating with a Foreign Faculty Member Regarding a Research Proposal}
I imagine that this includes the Fulbright and, to a greater extent, the Churchill. This summer, I am busy with rehabilitation and other academic things. For this reason, I believe that these things may not be for me right now.

\subsection{Study Abroad Programs that Just Look at Academic Excellence as a Criterion}
I imagine that this just includes the Clarendon. What is your advice for applying to this one?

\section{timeline}
in your experience, what is the general timeline for all these application processes that I should be aware of right now?
\end{document}
